### 2021-01-23 to_ams_editor_1.tex
first AMS version and letter to editor

\documentclass{ucrmath}
\signature{John Baez}
\address{Department of Mathematics \\ University of California \\ Riverside, CA}

\begin{document}

\begin{letter}{}
\opening{Dear Henri Darmon:}

This is a paper in the field of applied category theory.   It proves that two constructions of a double category of cospans equipped with extra structure are equivalent under certain conditions, and it illustrates the applications of these constructions in a number of fields.   I'm submitting it to you because the AMS website says you're the editor for ``unclassified'' papers at \textsl{Transactions of the American Mathematical Society}, and this paper doesn't fit squarely in the other classifications.

\closing{Sincerely,}

\end{letter}

\end{document}